Extracting Angular Observables without a Likelihood\\and Applications to Rare Decays

### 2014-07-27 draft.tex
technical commit

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[svgnames]{xcolor}
\usepackage{hepparticles}
\usepackage{hepnicenames}
\usepackage{hepunits}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{epstopdf}
%% Shortcuts %%
\newcommand{\rmdx}[1]{\mbox{d} #1 \,} % differential
\newcommand{\refapp}[1]{app.~(\ref{app:#1})}
\newcommand{\refeq}[1]{eq.~(\ref{eq:#1})}
\newcommand{\refsec}[1]{sec.~(\ref{sec:#1})}
\newcommand{\reftab}[1]{tab.~(\ref{tab:#1})}
\newcommand{\ie}{\textit{i.e.}}
\newcommand{\nuvec}{\vec{\nu}}
\newcommand{\thvec}{\vec{\vartheta}}
\let\oldtheta\theta
\renewcommand{\theta}{\vartheta}
%\newcommand{\GeV}{\text{GeV}}
\newcommand{\order}[1]{\mathcal{O}\left({#1}\right)}
\let\eps\varepsilon
\newcommand{\aver}[1]{\langle #1 \rangle}
\newcommand{\est}[1]{\widehat{#1}}
\newcommand{\wwhat}[1]{\widehat{\widehat{#1}}}
\DeclareMathOperator{\cov}{Cov}
%\DeclareMathOperator{\skew}{Skew}
\DeclareMathOperator{\kurt}{Kurt}

@@

\begin{abstract}
In this letter we study how to obtain a set of angular observables
$\{S_i\}$ that arise in a generic multi-body process without the need
to carry out a likelihood fit of an angular distribution to the
measured events. Instead, we present a method that only relies on
orthogonality of angular functions, and estimation of integrals
by means of Monte Carlo techniques. Beyond providing estimators
for central values and uncertainties, we provide means to determine
the correlation between all angular observables. We show that detector
acceptance effects can be accounted for in the analysis.
\end{abstract}

@@

We continue with basic definitions that pertain to angular observables, and our
results in the subsequent sections. Let $\thvec = (\theta_1, \dots \theta_N)$ denote the set of all
angles, and let $\nuvec = (\nu_1, \dots \nu_M)$ denote the set of all
other non-angular kinematic variables needed to fully specify the
final state of the process under study. For example, $\nuvec$ may
include invariant masses or center-of-mass energies. We define an
angular observable $S_i$ as a coefficient in the probability density
function (PDF) of the process by means of
\begin{align}
    \label{eq:def-P}
    P(\nuvec, \thvec) \equiv \sum_i S_i(\nuvec) \times f_i(\thvec)\,.
\end{align}
Here, the dependence on the $n$ decay angles $\thvec$ has been
explicitly factored out in terms of the angular functions
$f_i(\thvec)$, which we assume to fulfill the orthonormality relations
\begin{equation}
    \label{eq:def-ortho-rel}
    \int_\Omega \rmdx{^N \theta} f_i(\thvec) f_j(\thvec)  = \delta_{ij}\,,
\end{equation}
with $\Omega$ representing the full angular phase space relevant to the process.
For the purpose of this letter, however, it suffices that the system
of angular functions $f_i(\thvec)$ can be transformed into an
orthonormal basis. The transformation needs to be worked out on a
case-by-case basis. For a selection of $B$ decays which we consider the most interesting
we list the transformations in a series of appendices \ref{app:btokll} through
\ref{app:lambdabtolambdall}.\\

For particle decays, $P$ is generally expressed in terms of the fully differential decay width,
\begin{align}
    \label{eq:def-P-decay}
    P(\nuvec, \thvec) \equiv \frac{1}{\Gamma}\frac{\rmdx{^{N+M}\Gamma}}{\rmdx{\nu_1} \dots \rmdx{\nu_M} \rmdx{\theta_1} \dots \rmdx{\theta_N}}\,,
\end{align}
where $\Gamma$ is the total decay width. For a scattering process, one can similarly use
\begin{align}
    \label{eq:def-P-scattering}
    P(\nuvec, \thvec) \equiv \frac{1}{\sigma}\frac{\rmdx{^{N+M}\sigma}}{\rmdx{\nu_1} \dots \rmdx{\nu_M} \rmdx{\theta_1} \dots \rmdx{\theta_N}}\,,
\end{align}
where the total cross section $\sigma$ is used for the
normalization. Since the determination of the total decay width or
total cross section can be quite difficult, we emphasize that
different normalizations for $P$ can be used.  For instance, the total
decay width (or cross section) of the process of interest can be
replaced by the corresponding quantity of a control-channel
process. This change of normalization is equivalent to a linear rescaling of
the angular observables $\lbrace S_i\rbrace$; thus ratios or similar suitable combinations
of the angular observables are invariant.\\

@@

The orthonormality relations \refeq{def-ortho-rel} imply that a single angular observable $S_i$
can be projected out of the full PDF $P$ by means of
\begin{equation}
    \label{eq:det-Pi-analytical}
    S_i(\nuvec) = \int_{\Omega} \rmdx{^N \theta} \hat{f}_i(\thvec) P(\nuvec, \thvec)\,.
\end{equation}
where $\lbrace \hat{f}_i \rbrace$ denotes a \emph{dual basis} of angular function. In all generality
the bases $\lbrace f_i \rbrace$ and $\lbrace \hat{f}_i \rbrace$ are not identical. This is the
case for our selection of applications in appendices \ref{app:btokll} through \ref{app:lambdabtolambdall}.\\

It is sensible to refer to the angular observable $S_i$ as the \emph{$\hat{f}_i$-moment} of the PDF $P$.
We emphasize that a relation of type \refeq{det-Pi-analytical} holds for any combination of
scalar product and orthonormal basis of angular functions $\lbrace f_i \rbrace$; \ie, there is no unique
basis of angular functions. For the proof we refer to ref. \cite{Dighe:1998vk}.\\
In addition, integration over the non-angular variables yields
\begin{equation}
    \langle S_i\rangle
    \equiv \int \rmdx{^M \nu} S_i(\nuvec)
    = \int \rmdx{^M \nu} \left[\int_{\Omega} \rmdx{^N \theta}\hat{f}_i(\thvec) P(\nuvec,\thvec) \right].
\end{equation}
The remainder of this section describes the method of moments, in which we replace the
analytical integration \refeq{det-Pi-analytical} by Monte Carlo (MC) estimators.\\

The central tenet of MC integration is the fact that the expectation value $E_P[g]$ of some function
$g(x)$ under the probability density $P(x)$,
\begin{equation}
    E_P[g] \equiv \int \rmdx{x} g(x) P(x)
\end{equation}
can be approximated \cite{MCSM-2004} by an MC estimator $\est{E_P[g]}$
\begin{equation}
    \label{eq:mc-id}
    E_P[g] \to \widehat{E_P[g]} \equiv \frac{1}{K} \sum_{k=1} g(x^{(k)})
\end{equation}
due to the strong law of large numbers for $K \to \infty$ assuming
that the variates $x^{(k)}$, $k = 1, \dots, K$ are distributed as
$P$; i.e.,
\begin{equation}
    x^{(k)} \sim P\,.
\end{equation}
Throughout this letter we denote all MC estimators with a wide hat.\\

Application of \refeq{mc-id} then yields
\begin{equation}
    \langle S_i\rangle \to \widehat{\langle S_i\rangle} = \frac{1}{K} \sum_{k=1}^{K} \hat{f}_i(x^{(k)})\,.
\end{equation}
It is often of interest to obtain the $\nuvec$- observables integrated
over certain ranges of $\nuvec$; i.e., binned measurements of $\langle
S_i\rangle$.
Therefore, we introduce the bin-integrated quantity
\begin{align}
    \langle S_i\rangle_{\vec{a},\vec{b}}
    & = \int_{\vec{a}}^{\vec{b}} \rmdx{^M \nu} S_i(\nuvec)\\
    & = \int_{\vec{a}}^{\vec{b}} \rmdx{^M \nu} \left[\int_{\Omega} \rmdx{^N\theta} \hat{f}_i(\thvec) P(\nuvec,\thvec)  \right]\\
    & = \int \rmdx{^M \nu} \left[\int_{\Omega} \rmdx{^N\theta} \hat{f}_i(\thvec) P(\nuvec,\thvec)
        \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,
        \right],
\end{align}
where the argument of the indicator function $ \mathbf{1}(\vec{a} \le
\nuvec \le \vec{b})$ is to be interpreted componentwise.
Application of \refeq{mc-id} immediately yields
\begin{equation}
    \label{eq:bin-importance}
    \widehat{\langle S_i\rangle_{\vec{a},\vec{b}}}
    = \frac{1}{K} \sum_{k=1}^{K} \hat{f}_i(x^{(k)})         \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,.
\end{equation}
This basically reduces the estimation of
$\aver{S_i}_{\vec{a},\vec{b}}$ to a (weighted) counting experiment. In
the limit $N \to \infty$, the random vector $( \dots,
\aver{S_i}_{\vec{a},\vec{b}}, \dots)$ follows a multivariate Gaussian
distribution \citeneeded\fred{Robert,Casella?}
with the mean vector estimated by \refeq{bin-importance}, and the covariance
estimated as
\begin{equation}
    \cov[\aver{S_i},\aver{S_j}]_{\vec{a},\vec{b}} \to \est{\cov}[\aver{S}_i, \aver{S}_j]_{\vec{a},\vec{b}}
        = \frac{1}{K - 1} \sum_{k=1}^{K} \big[\hat{f}_i(x^{(k)}) - \est{\langle S_i\rangle_{\vec{a},\vec{b}}}\big]\,\big[\hat{f}_j(x^{(k)}) - \widehat{\langle S_j\rangle_{\vec{a},\vec{b}}}\big]
        \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,.
\end{equation}
We would like to emphasize that the sample covariance should rapidly converge towards the true covariance matrix.
We propose to declare convergence once estimators for skewness and kurtosis are sufficiently small,
\begin{equation}
    \est{\operatorname{Skew}}[\aver{S_i},\aver{S_j}] \ll 1\,\,\text{ and }\,\,\est{\kurt}[\aver{S_i},\aver{S_j}] \ll 1\,.
\end{equation}

@@

\begin{figure}
    \includegraphics[width=.6\textwidth]{fig-topology.pdf}
    \caption{Decay topology for decays $B\to P_1 P_2 \ell_1 \bar\ell_2$. \label{fig:topology}}
\end{figure}

@@

As before, we assume a PDF $P$ can be decomposed into $N$ parts with angular observables $\lbrace S_k\rbrace$,
\begin{equation}
    P \equiv P(\vec\nu,\vec\theta | \lbrace S_k \rbrace)\,.
\end{equation}
We denote the estimator of an
observable $S_k$ as $\est{S_k}$, which is affected by the detector acceptance. We assume now that there is a linear relation
between the true angular observables $\lbrace S_k^\text{true}\rbrace$, and the estimators $\lbrace \est{S_k}\rbrace$,
\begin{equation}
    \label{eq:linear-assumption}
    \est{S_i} = \sum_{i = 1}^N M_{i j} S_j^\text{true}\,.
\end{equation}
This assumption is based on a smooth angular acceptance function. While this will not be true for the laboratory frame,
smoothness may be recovered in boosted frame, e.g. the rest frame of the decaying particle.
In that case, the matrix $M_{ij}$ can be fully determined through the following steps.
\begin{enumerate}
    \item For a fixed $j$, let $S_j^\text{true} = 1$, and $S_k^\text{true} = 0\,\forall k\neq j$.
    \item Draw simulated events $\lbrace x_t^{(k)}\rbrace = \lbrace(\vec{\nu}^{(k)}_t,\vec\theta^{(k)}_t)\rbrace$ (dubbed ``true'') from the PDF $P_j$,
        \begin{equation}
            P_j \equiv P(\vec\nu,\vec\theta | \lbrace S_j = 1, S_{k\neq j} = 0\rbrace)\,.
        \end{equation}
    \item Use the true events as input for the MC simulation of the detector, in order to obtain the detector
        events $\lbrace x^{(k)}_d\rbrace = \lbrace(\vec{\nu}^{(k)}_d,\vec\theta^{(k)}_d)\rbrace$.
    \item Use the method of moments to determine the uncorrected observables $\lbrace \est{S_k} \rbrace$, based on the detector
        events $\lbrace x^{(k)}_d\rbrace$.
    \item By virtue of \refeq{linear-assumption} and step 1, the $j$-th column of $M$ has been determined via
        \begin{equation}
            M_{ij} = \est{S_i}\,.
        \end{equation}
        \fred{I wonder if a small sample size also leads to artificial
          nonzero off-diagonal elements of $M$}
    \item Repeat the above steps for all values of $j$ to determine all columns of $M$.
\end{enumerate}

@@

We emphasize that this methods hinges on the assumption in \refeq{linear-assumption}, We thus suggest to explicitly
test the assumption. One way of testing involves repeating the previous procedure
with $S_j^\text{true} = 1/x$. If the lineary assumption holds, one should find
\begin{equation}
    M_{ij} = x \est{S_i}\,,
\end{equation}
for arbitrary values of $x > 1$.

@@

The PDF for the decay $\bar{B}\to\bar{K}\ell^+\ell^-$ has been calculated for the most
complete basis of dimension-six $b\to s \ell^+\ell^-$ operators. It reads \cite{Bobeth:2007dw,Bobeth:2012vn}
\begin{equation}
    P(q^2, \cos\theta_1) = \frac{1}{\Gamma} \frac{\rmdx{^4\Gamma}}{\rmdx{q^2} \rmdx{\cos\theta_1}} = \frac{a(q^2)}{\Gamma} + \frac{b(q^2)}{\Gamma} \cos\theta_1 + \frac{c(q^2)}{\Gamma} \cos^2\theta_1 = \sum_i P_i f_i(\cos\theta_1)\,,
\end{equation}
where the basis of angular functions $\lbrace f_i\rbrace$ reads
\begin{equation}
\begin{aligned}
    f_1 & = 1\,, &
    f_2 & = \cos\theta_1\,, &
    f_2 & = \cos^2\theta_1\,.
\end{aligned}
\end{equation}
Here we denote the dilepton mass squared as $q^2$, and the dilepton helicity angle as $\theta_1 \equiv \theta_{\ell}$. We transform the PDF from a $\cos\theta_1$-dependence
to a $\theta_1$-dependence,
\begin{equation}
    P(q^2, \theta_1) = \sum_i P_i f_i(\cos\theta_1) \sin \theta_1\,.
\end{equation}
where we define $P_{\lbrace 1,2,3\rbrace} \equiv \lbrace a, b, c\rbrace / \Gamma$.\\

The dual basis $\lbrace \hat{f}_i\rbrace$ reads
\begin{equation}
\begin{aligned}
    \hat{f}_1 & = \frac{9}{8} - \frac{15}{8}\cos^2\theta_1\,, &
    \hat{f}_2 & = \frac{3}{2}\cos\theta_1\,, &
    \hat{f}_3 & = -\frac{15}{8} + \frac{45}{8}\cos^2\theta_1\,,
\end{aligned}
\end{equation}
such that
\begin{equation}
    \int_0^\pi \rmdx{\theta_1} \hat{f}_i(\theta_1) P(q^2, \theta_1) = S_i(q^2)\,.
\end{equation}

>>>>

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[svgnames]{xcolor}
\usepackage{hepparticles}
\usepackage{hepnicenames}
\usepackage{hepunits}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{epstopdf}
%% Shortcuts %%
\newcommand{\dd}{\text{d}}
\newcommand{\refapp}[1]{app.~(\ref{app:#1})}
\newcommand{\refeq}[1]{eq.~(\ref{eq:#1})}
\newcommand{\refsec}[1]{sec.~(\ref{sec:#1})}
\newcommand{\reftab}[1]{tab.~(\ref{tab:#1})}
\newcommand{\ie}{\textit{i.e.}}
\newcommand{\nuvec}{\vec{\nu}}
\newcommand{\thvec}{\vec{\vartheta}}
\let\oldtheta\theta
\renewcommand{\theta}{\vartheta}
%\newcommand{\GeV}{\text{GeV}}
\newcommand{\order}[1]{\mathcal{O}\left({#1}\right)}
\let\eps\varepsilon
\newcommand{\aver}[1]{\langle #1 \rangle}
\newcommand{\est}[1]{\widehat{#1}}
\newcommand{\wwhat}[1]{\widehat{\widehat{#1}}}
\DeclareMathOperator{\cov}{Cov}
%\DeclareMathOperator{\skew}{Skew}
\DeclareMathOperator{\kurt}{Kurt}

@@

\begin{abstract}
In this letter we study how to obtain a set of angular observables
$\{P_i\}$ that arise in a generic multi-body process without the need
to carry out a likelihood fit of an angular distribution to the
measured events. Instead, we present a method that only relies on
orthogonality of angular functions, and estimation of integrals
by means of Monte Carlo techniques. Beyond providing estimators
for central values and uncertainties, we provide means to determine
the correlation between all angular observables. We show that detector
acceptance effects can be accounted for in the analysis.
\end{abstract}

@@

We continue with basic definitions that pertain to angular observables, and our
results in the subsequent sections. Let $\thvec = (\theta_1, \dots \theta_N)$ denote the set of all
angles, and let $\nuvec = (\nu_1, \dots \nu_M)$ denote the set of all
other non-angular kinematic variables needed to fully specify the
final state of the process under study. For example, $\nuvec$ may
include invariant masses or center-of-mass energies. We define an
angular observable $S_i$ as a coefficient in the probability density
function (PDF) of the process by means of
\begin{align}
    \label{eq:def-P}
    P(\nuvec, \thvec) \equiv \sum_i S_i(\nuvec) \times f_i(\thvec)\,.
\end{align}
Here, the dependence on the $n$ decay angles $\thvec$ has been
explicitly factored out in terms of the angular functions
$f_i(\thvec)$, which we assume to fulfill the orthonormality relations
\begin{equation}
    \label{eq:def-ortho-rel}
    \int_\Omega f_i(\thvec) f_j(\thvec) \dd^N \theta = \delta_{ij}\,,
\end{equation}
with $\Omega$ representing the full angular phase space relevant to the process.
For the purpose of this letter, however, it suffices that the system
of angular functions $f_i(\thvec)$ can be transformed into an
orthonormal basis. The transformation needs to be worked out on a
case-by-case basis. For a selection of $B$ decays which we consider the most interesting
we list the transformations in a series of appendices \ref{app:btokll} through
\ref{app:lambdabtolambdall}.\\

For particle decays, $P$ is generally expressed in terms of the fully differential decay width,
\begin{align}
    \label{eq:def-P-decay}
    P(\nuvec, \thvec) \equiv \frac{1}{\Gamma}\frac{\dd^{N+M}\Gamma}{\dd \nu_1 \dots \dd \nu_M\, \dd \theta_1 \dots \theta_N}\,,
\end{align}
where $\Gamma$ is the total decay width. For a scattering process, one can similarly use
\begin{align}
    \label{eq:def-P-scattering}
    P(\nuvec, \thvec) \equiv \frac{1}{\sigma}\frac{\dd^{N+M}\sigma}{\dd \nu_1 \dots \dd \nu_M\, \dd \theta_1 \dots \theta_N}\,,
\end{align}
where the total cross section $\sigma$ is used for the
normalization. Since the determination of the total decay width or
total cross section can be quite difficult, we emphasize that
different normalizations for $P$ can be used.  For instance, the total
decay width (or cross section) of the process of interest can be
replaced by the corresponding quantity of a control-channel
process. This change of normalization is equivalent to a linear rescaling of
the angular observables $\lbrace S_i\rbrace$; thus ratios or similar suitable combinations
of the angular observables are invariant.\\

@@

The orthonormality relations \refeq{def-ortho-rel} imply that a single angular observable $S_i$
can be projected out of the full PDF $P$ by means of
\begin{equation}
    \label{eq:det-Pi-analytical}
    S_i(\nuvec) = \int_{\Omega} \hat{f}_i(\thvec) P(\nuvec, \thvec) \dd^N \theta\,.
\end{equation}
where $\lbrace \hat{f}_i \rbrace$ denotes a \emph{dual basis} of angular function. In all generality
the bases $\lbrace f_i \rbrace$ and $\lbrace \hat{f}_i \rbrace$ are not identical. This is the
case for our selection of applications in appendices \ref{app:btokll} through \ref{app:lambdabtolambdall}.\\

It is sensible to refer to the angular observable $S_i$ as the \emph{$\hat{f}_i$-moment} of the PDF $P$.
We emphasize that a relation of type \refeq{det-Pi-analytical} holds for any combination of
scalar product and orthonormal basis of angular functions $\lbrace f_i \rbrace$; \ie, there is no unique
basis of angular functions. For the proof we refer to ref. \cite{Dighe:1998vk}.\\
In addition, integration over the non-angular variables yields
\begin{equation}
    \langle S_i\rangle
    \equiv \int S_i(\nuvec) \dd^M \nu
    = \int \left[\int_{\Omega} \hat{f}_i(\thvec) P(\nuvec,\thvec) \dd^N\theta\right]\dd^M \nu\,.
\end{equation}
The remainder of this section describes the method of moments, in which we replace the
analytical integration \refeq{det-Pi-analytical} by Monte Carlo (MC) estimators.\\

The central tenet of MC integration is the fact that the expectation value $E_P[g]$ of some function
$g(x)$ under the probability density $P(x)$,
\begin{equation}
    E_P[g] \equiv \int g(x) P(x) \dd x
\end{equation}
can be approximated \cite{MCSM-2004} by an MC estimator $\est{E_P[g]}$
\begin{equation}
    \label{eq:mc-id}
    E_P[g] \to \widehat{E_P[g]} \equiv \frac{1}{K} \sum_{k=1} g(x^{(k)})
\end{equation}
due to the strong law of large numbers for $K \to \infty$ assuming
that the variates $x^{(k)}$, $k = 1, \dots, K$ are distributed as
$P$; i.e.,
\begin{equation}
    x^{(k)} \sim P\,.
\end{equation}
Throughout this letter we denote all MC estimators with a wide hat.\\

Application of \refeq{mc-id} then yields
\begin{equation}
    \langle S_i\rangle \to \widehat{\langle S_i\rangle} = \frac{1}{K} \sum_{k=1}^{K} \hat{f}_i(x^{(k)})\,.
\end{equation}
It is often of interest to obtain the $\nuvec$- observables integrated
over certain ranges of $\nuvec$; i.e., binned measurements of $\langle
S_i\rangle$.
Therefore, we introduce the bin-integrated quantity
\begin{align}
    \langle S_i\rangle_{\vec{a},\vec{b}}
    & = \int_{\vec{a}}^{\vec{b}} S_i(\nuvec) \dd^M \nu\\
    & = \int_{\vec{a}}^{\vec{b}} \left[\int_{\Omega} \hat{f}_i(\thvec) P(\nuvec,\thvec) \dd^N\theta\right]\dd^M \nu\\
    & = \int \left[\int_{\Omega} \hat{f}_i(\thvec) P(\nuvec,\thvec)
        \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,
        \dd^N\theta\right]\dd^M \nu\,,
\end{align}
where the argument of the indicator function $ \mathbf{1}(\vec{a} \le
\nuvec \le \vec{b})$ is to be interpreted componentwise.
Application of \refeq{mc-id} immediately yields
\begin{equation}
    \label{eq:bin-importance}
    \widehat{\langle S_i\rangle_{\vec{a},\vec{b}}}
    = \frac{1}{K} \sum_{k=1}^{K} \hat{f}_i(x^{(k)})         \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,.
\end{equation}
This basically reduces the estimation of
$\aver{S_i}_{\vec{a},\vec{b}}$ to a (weighted) counting experiment. In
the limit $N \to \infty$, the random vector $( \dots,
\aver{S_i}_{\vec{a},\vec{b}}, \dots)$ follows a multivariate Gaussian
distribution \citeneeded\fred{Robert,Casella?}
with the mean vector estimated by \refeq{bin-importance}, and the covariance
estimated as
\begin{equation}
    \cov[\aver{S_i},\aver{S_j}]_{\vec{a},\vec{b}} \to \est{\cov}[\aver{S}_i, \aver{S}_j]_{\vec{a},\vec{b}}
        = \frac{1}{K - 1} \sum_{k=1}^{K} \big[\hat{f}_i(x^{(k)}) - \est{\langle S_i\rangle_{\vec{a},\vec{b}}}\big]\,\big[\hat{f}_j(x^{(k)}) - \widehat{\langle S_j\rangle_{\vec{a},\vec{b}}}\big]
        \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,.
\end{equation}
We would like to emphasize that the sample covariance should rapidly converge towards the true covariance matrix.
We propose to declare convergence once estimators for skewness and kurtosis are sufficiently small,
\begin{equation}
    \est{\operatorname{Skew}}[\aver{S_i},\aver{S_j}] \ll 1\,\,\text{ and }\,\,\est{\kurt}[\aver{S_i},\aver{S_j}] \ll 1\,.
\end{equation}

@@

\fred{Paragraph is very specific to one physics case. Needs a drawing}

@@

As before, we assume a PDF $P$ can be decomposed into $N$ parts with angular observables $\lbrace S_k\rbrace$,
\begin{equation}
    P \equiv P(\vec\nu,\vec\theta | \lbrace S_k \rbrace)\,.
\end{equation}
We denote the estimator of an
observable $S_k$ as $\est{S_k}$, which is affected by the detector acceptance. We assume now that there is a linear relation
between the true angular observables $\lbrace S_k^\text{true}\rbrace$, and the estimators $\lbrace \est{S_k}\rbrace$,
\fred{Do we need $S_k^\text{true}$? Isn't it just $S_k$?}
\begin{equation}
    \label{eq:linear-assumption}
    \est{S_i} = \sum_{i = 1}^N M_{i j} S_j^\text{true}\,.
\end{equation}
In that case, the matrix $M_{ij}$ can be fully determined through the following steps.
\begin{enumerate}
    \item For a fixed $j$, let $S_j^\text{true} = 1$, and $S_k^\text{true} = 0\,\forall k\neq j$.
    \item Draw simulated events $\lbrace x_t^{(k)}\rbrace = \lbrace(\vec{\nu}^{(k)}_t,\vec\theta^{(k)}_t)\rbrace$ (dubbed ``true'') from the PDF $P_j$,
        \begin{equation}
            P_j \equiv P(\vec\nu,\vec\theta | \lbrace S_j = 1, S_{k\neq j} = 0\rbrace)\,.
        \end{equation}
    \item Use the true events as input for the MC simulation of the detector, in order to obtain the detector
        events $\lbrace x^{(k)}_d\rbrace = \lbrace(\vec{\nu}^{(k)}_d,\vec\theta^{(k)}_d)\rbrace$.
    \item Use the method of moments to determine the uncorrected observables $\lbrace \est{S_k} \rbrace$, based on the detector
        events $\lbrace x^{(k)}_d\rbrace$.
    \item By virtue of \refeq{linear-assumption} and step 1, the $j$-th column of $M$ has been determined via
        \begin{equation}
            M_{ij} = \est{S_i}\,.
        \end{equation}
        \fred{I wonder if a small sample size also leads to artificial
          nonzero off-diagonal elements of $M$}
    \item Repeat the above steps for all values of $j$ to determine all columns of $M$.
\end{enumerate}

@@

We emphasize that this methods hinges on the assumption in \refeq{linear-assumption}, which might not be fulfilled in general.
In particular, if detector effects are small, of it they can be reduced through other means, our approach is expected to work.
The effects can be considered small if
\begin{equation}
    ||M - \mathbf{1}||_2 \ll N\,.
\end{equation}
Here $||\cdot||_2$ denotes the Frobenius norm.

\danny{$@$Marcin: How well does this work? Is \refeq{linear-assumption} fulfilled in your cases?}

\vspace{\bigskipamount}
\hrule
\danny{Should we remove the remainder of this section?}
\fred{Comment that in this section $\vec\nu$ and $\vec\theta$ are used after ``correction'' has been applied.}
Thanks to accurate detector simulations and control studies, the experimental analyses have become
remarkably good at removing detector acceptance effects. In this section we therefore assume that
there are only a small, residual acceptance effects left that affect the events. In particular, we
assume that the latter can be modeled as a modified probability density,
\begin{equation}
    \label{eq:PDFwithDE}
    P(\nuvec,\thvec) \to P(\nuvec,\thvec) \times \Omega(\thvec),
\end{equation}
where $\Omega$ is independent of the non-angular kinematics $\nuvec$. The latter assumption is only
is invalid in general. However, if residual angular acceptance only weakly depends on $\vec{\nu}$, one
can always choose appropriate bin $\vec{\nu}$ so that $\Omega$ is approximately constant with respect
to $\vec{\nu}$. For such a case we parametrize
\begin{equation}
    \Omega_{m,l}(\thvec) = \prod_{i=1}^{m} \Big[1 + \sum_{j=1}^l \eps_i^{(j)} p_j(\cos\theta_i)\big]
\end{equation}
where $p_j(\cos\theta_i)$ denotes the $j$th Legendre polynomial. Since we only consider residual
acceptance effects, we expect $\eps_i^{(j)} \ll 1$. \footnote{\danny{what are reasonable values/upper limits on the $\eps$?}}\\

We are aware that setting a bin width that fulfills the aforementioned criterion can be at odds with
other requirements. For instance, in exclusive $b\to s\ell^+\ell^-$ decays, the bin width for
$\vec{\nu} = q^2$ needs to cover a sufficiently large part of the phase space beyond the open charm
threshold. We emphasize that the outcome of an analysis as proposed here can easily combine several
bins by means of a weighted average.\\

\paragraph{Uniangular Distribution}
For definiteness and clarity, we consider first an uni-angular distribution $P(\theta)$.
In this case we can parametrize the PDF $P$ as
\begin{equation}
    P(\theta) = S_1 + S_2 \cos\theta + S_2 \cos^2\theta\,,
\end{equation}
which is a parametrization that has been used in the literature for $\bar{B}\to\bar{K}\ell^+\ell^-$ decays \cite{Bobeth:2007dw}.
While this system of angular functions is not orthonormal, it can easily be orthonormalized. For the basis of angular
functions, see \refapp{btokll}.
Considering detector acceptance effects up to second order in $\cos\theta$, the function $\Omega(\vec\theta)$ reads
\begin{equation}
    \Omega_{1,2}(\theta_1) = 1 + \eps_{1}^{(1)} \cos\theta + \frac{\eps_{1}^{(2)}}{2}\big(3 \cos^2\theta - 1\big)\qquad\eps_{1}^{(i)} = \order{\eps}\,.
\end{equation}
Inclusion of $\Omega_{1,2}$ as part of the PDF as proposed in \refeq{PDFwithDE} modifies the MC estimators for the angular observables.
We find
\begin{equation}
\begin{aligned}
    \wwhat{S_1} & = \Big(1 + \frac{\eps_{1}^{(2)}}{2}\Big) \est{S_1} + \frac{9 \eps_{1}^{(2)}}{70} \est{S_3}\\
    \wwhat{S_2} & = \Big(1 - \frac{2 \eps_{1}^{(2)}}{5}\Big) \est{S_2} - \eps_{1}^{(1)} \est{S_1} - \frac{3\eps_{1}^{(1)}}{5} \est{P_3}\\
    \wwhat{S_3} & = \Big(1 - \frac{11\eps_{1}^{(2)}}{14}\Big)\est{S_3} - \frac{3\eps_{1}^{(2)}}{2} \est{S_1} - \eps_{1}^{(1)} \est{P_2}\,.
\end{aligned}
\end{equation}
Here $\lbrace \est{S_i}\rbrace$ are the uncorrected estimators, while $\lbrace \wwhat{S_i}\rbrace$ denote the
acceptance-corrected estimators of the angular observables.\\

\paragraph{Triangular Distribution}

\begin{table*}
    \resizebox{.99\textwidth}{!}{\ensuremath{\input{tab-btokstarll-1to6.tex}}}\\[2.5\medskipamount]
    \resizebox{.99\textwidth}{!}{\ensuremath{\input{tab-btokstarll-7to12.tex}}}\\[2.5\medskipamount]
    \resizebox{.99\textwidth}{!}{\ensuremath{\input{tab-btokstarll-13to18.tex}}}
    \caption{The matrix $N_{ij}$ that appears in \refeq{eff-corrected-est-btokstarll}. Due to spatial
        constraints, we split its display into columns $j=1s,1c,1i,2s,2c,2i$ (top),
        $j=3,4,4i,5,5i,6s$ (middle) and $j=6c,7,7i,8,8i,9$ (bottom).
        \label{tab:eff-corrected-est-btokstarll}
    }
\end{table*}

As an example for a triangular distribution we use the decay $\bar{B}\to\bar{K}^*(\to \bar{K}\pi)\ell^+\ell^-$.
A parametrization of its angular distribution can be found in \refapp{btokstarll}.\\

For this type of distribution, we expand $\Omega_3$ to linear order in $\eps$,
\begin{equation}
    \Omega_{3,2}(\theta_1,\theta_2,\theta_3) = 1 + \sum_{i=1}^3 \sum_{j=1}^2 \eps_i^{(j)} p_j(\cos\theta_i) + \order{\eps^2}\,.
\end{equation}
This leads to a matrix-valued equation that relates the effiency-corrected estimators $\wwhat{S}_i$ to the uncorrected
estimators $\est{S}_j$,
\begin{equation}
    \label{eq:eff-corrected-est-btokstarll}
    \wwhat{S}_i = \sum_{j=1\text{s},\dots\,9} \big(\delta_{ij} + N_{ij}\big) \est{S}_j\,.
\end{equation}
We provide the matrix $N$ in \reftab{eff-corrected-est-btokstarll}.\\
\hrule

@@

The PDF for the decay $\bar{B}\to\bar{K}\ell^+\ell^-$ has been calculated for the most
complete basis of dimension-six $b\to s \ell^+\ell^-$ operators. It reads \cite{Bobeth:2007dw,Bobeth:2012vn}
\begin{equation}
    P(q^2, \cos\theta_1) = \frac{1}{\Gamma} \frac{\dd^4\Gamma}{\dd q^2\,\dd \cos\theta_1} = \frac{a(q^2)}{\Gamma} + \frac{b(q^2)}{\Gamma} \cos\theta_1 + \frac{c(q^2)}{\Gamma} \cos^2\theta_1 = \sum_i P_i f_i(\cos\theta_1)\,,
\end{equation}
where the basis of angular functions $\lbrace f_i\rbrace$ reads
\begin{equation}
\begin{aligned}
    f_1 & = 1\,, &
    f_2 & = \cos\theta_1\,, &
    f_2 & = \cos^2\theta_1\,.
\end{aligned}
\end{equation}
Here we denote the dilepton mass squared as $q^2$, and the dilepton helicity angle as $\theta_1 \equiv \theta_{\ell}$. We transform the PDF from a $\cos\theta_1$-dependence
to a $\theta_1$-dependence,
\begin{equation}
    P(q^2, \theta_1) = \sum_i P_i f_i(\cos\theta_1) \sin \theta_1\,.
\end{equation}
where we define $P_{\lbrace 1,2,3\rbrace} \equiv \lbrace a, b, c\rbrace / \Gamma$.\\

The dual basis $\lbrace \hat{f}_i\rbrace$ reads
\begin{equation}
\begin{aligned}
    \hat{f}_1 & = \frac{9}{8} - \frac{15}{8}\cos^2\theta_1\,, &
    \hat{f}_2 & = \frac{3}{2}\cos\theta_1\,, &
    \hat{f}_3 & = -\frac{15}{8} + \frac{45}{8}\cos^2\theta_1\,,
\end{aligned}
\end{equation}
such that
\begin{equation}
    \int_0^\pi \dd \theta_1\hat{f}_i(\theta_1) P(q^2, \theta_1) = S_i(q^2)\,.
\end{equation}
